Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer

### 2019-02-25 Paper/Paper.tex
Beginning to Address Qing's Comments

\section*{Abstract}
%% Basic Introduction (1-2 sentences)
Preceding tumorigenesis and throughout the lifetime of a tumor, the immune system is engaged in complex sets of interactions in and around the tumor.
%% More Detailed Background (2-3 sentences)
These are regulated by the tumor microenvironment (TME) in myriad ways. Moreover,  epithelial cell plasticity impacts disease incidence and progression via processes of epithelial-to-mesenchymal transition (EMT).
In turn, both the tumor and the immune system exert substantial influence over the TME in competition to influence the fate of the host.
%% General Problem (1 sentence)
Determining the factors and mechanisms underpinning these complex interactions remains an important task for cancer immunotherapy.
%% Summarizing Main Result (1 sentence)
%Here, we seek to understand how interplay between the tumor, the immune system, and EMT can determine the outcome of cancer.
%% The Main Result (2-3 sentences)
We developed a multiscale agent-based model to describe the interplay between three components: the tumor (or pre-tumor cells), the immune system, and EMT.
The model revealed a non-monotonic relationship between the cell cycle times associated with the mesenchymal phenotype and the probability of tumorigenesis, identifying a critical growth rate that maximizes cancer-free survival. In comparison, increasing the immune-evasiveness of the mesenchymal phenotypes always led to poorer prognosis. Moreover, encouraging
%\tcr
the TME
towards a pro-EMT state also benefits the tumor.
%% Putting into a More General Context (1-2 sentences)
Together, these results indicate that the joint regulation of the TME by the immune system and the tumor itself provides possible targets for future cancer therapies, in particular cancers such as pancreatic cancer in which EMT is speculated to play a role.

@@

\section{Introduction}
Immunotherapy is slowly revolutionizing cancer therapy. New breakthroughs in the field begin to produce new therapies that have had a significant impact on patient health and survival\cite{pardoll2012blockade}\cite{restifo2012adoptive}.
% A More examples? Reviews or non-reviews? Explain some of the examples?
%AM I will add a few 
%There are myriad untapped avenues by which scientists can further explore the role of the immune system in regulating and combatting cancer;
In addition to direct cell-cell interactions between immune and cancer cells, the means by which immune cells interact with and modulate the tumor microenvironment (TME) is an important research area. Immune cells can be understood as one constituent of the TME: responding to the environment's inflammatory signals, fighting the mutated cells present, and then restoring homeostasis in that local context\cite{de2006paradoxical}.
Understanding each step in this immune cascade in the context of the TME will be necessary in order to fully explore and exploit all options of immunotherapy.

@@

The tumor interacts with the immune system by initiating the sequence of events that results in the activation of the adaptive immune response.
It accomplishes this by presenting antigens that innate immune cells recognize and carry to lymph nodes to activate, among other things, T cells.
Also, tumor cells engage in certain cellular processes that shape the TME and thus impact the immune system.
In particular, tumor cells release transforming growth factor beta (TGF-$\beta$) that creates a tumor-supportive environment by shifting the immune balance to being immunosuppressive via enhancing activation of Treg cells.

@@

Epithelial-to-mesenchymal transition (EMT) describes the (reversible) cell plasticity process by which epithelial cells transition into mesenchymal cells.
While the epithelial cells often stay bound together in layers, mesenchymal cells have gained motility and even stem-like properties\cite{nieto2016emt}.
Recent research by Hong has revealed the intriguing possibility that there are four stable equilibria along the EMT spectrum with each state having its own unique properties\cite{hong2015ovol2}.
They have thus become a subject of great interest in cancer research having been connected to tumor initiation, progression, and metastasis\cite{nieto2016emt}.
There are many other characteristics distinctive of mesenchymal cells and still much research being done in further quantifying these and determining others.
Of interest to us, the literature reveals two qualities of mesenchymal cells that are particularly relevant in the context of cancer and the immune system.
The first is that mesenchymal cells are less susceptible to immune clearance\cite{terry2017new}. 
%
The second is that mesenchymal cells proliferate at a slower rate. Ref
 %A Adam said he had a reference in mind for this.

The research on mesenchymal cells' immune evasiveness is quite conclusive with the research even going so far as establishing functional causes for this phenomenon.
As a cell is being targeted by cytotoxic immune cells for clearance, a physical connection between the two cells must be created in order to precipitate the lysing event.
This immunological synapse, however, is dependent upon the surface proteins of the target cell, and in the case of mesenchymal cells there is preliminary evidence that these surface markers are down-regulated in such a way that the immune synapse is more difficult to form\cite{terry2017new}.
There are also other structural changes that take place in mesenchymal cells that could contribute to this refractory nature towards immune clearance.
Regardless, the evidence currently demonstrates a ``slipperiness'' of mesenchymal cells in regards to immune lysing that in our model we will refer to as mesenchymal immune evasion, or MIE.

>>>>

\section*{Abstract}
%% Basic Introduction (1-2 sentences)
Preceding tumorigenesis and throughout the lifetime of a tumor, the immune system is engaged in complex sets of interactions in and around the tumor.
%% More Detailed Background (2-3 sentences)
These are regulated by the tumor microenvironment (TME) in myriad ways. Moreover,  epithelial cell plasticity impacts disease incidence and progression via processes of epithelial-to-mesenchymal transition (EMT).
In turn, both the tumor and the immune system exert substantial influence over the TME in competition to influence the fate of the host.
%% General Problem (1 sentence)
Determining the factors and mechanisms underpinning these complex interactions remains an important task for cancer immunotherapy.
%% Summarizing Main Result (1 sentence)
%Here, we seek to understand how interplay between the tumor, the immune system, and EMT can determine the outcome of cancer.
%% The Main Result (2-3 sentences)
We developed a multiscale agent-based model to describe the interplay between three components: the tumor (or pre-tumor cells), the immune system, and EMT.
The model revealed a \tcr{non-monotonic relationship} between the cell cycle times associated with the mesenchymal phenotype and the probability of tumorigenesis, identifying a critical growth rate that maximizes cancer-free survival. In comparison, increasing the immune-evasiveness of the mesenchymal phenotypes always led to poorer prognosis. Moreover, encouraging
%\tcr
the TME
towards a pro-EMT state also benefits the tumor.
%% Putting into a More General Context (1-2 sentences)
Together, these results indicate that the joint regulation of the TME by the immune system and the tumor itself provides possible targets for future cancer therapies, in particular cancers such as pancreatic cancer in which EMT is speculated to play a role.

@@

\section{Introduction}
Immunotherapy is slowly revolutionizing cancer therapy. New breakthroughs in the field begin to produce new therapies that have had a significant impact on patient health and survival\cite{pardoll2012blockade}\cite{restifo2012adoptive}.
% A More examples? Reviews or non-reviews? Explain some of the examples?
%AM I will add a few 
%There are myriad untapped avenues by which scientists can further explore the role of the immune system in regulating and combatting cancer;
In addition to direct cell-cell interactions between immune and cancer cells, the means by which immune cells interact with and modulate the tumor microenvironment (TME) is an important \tcr{research} area.
Immune cells can be understood as one constituent of the TME: responding to the environment's inflammatory signals, fighting the mutated cells present, and then restoring homeostasis in that local context\cite{de2006paradoxical}.
Understanding each step in this immune cascade in the context of the TME will be necessary in order to fully explore and exploit all options of immunotherapy.

@@

\tcr{Still need to address Qing's comments on this paragraph.}
The tumor interacts with the immune system by initiating the sequence of events that results in the activation of the adaptive immune response.
It accomplishes this by presenting antigens that innate immune cells recognize and carry to lymph nodes to activate, among other things, T cells.
Also, tumor cells engage in certain cellular processes that shape the TME and thus impact the immune system.
In particular, tumor cells release transforming growth factor beta (TGF-$\beta$) that creates a tumor-supportive environment by shifting the immune balance to being immunosuppressive via enhancing activation of Treg cells.

@@

Epithelial-to-mesenchymal transition (EMT) describes the (reversible) cell plasticity process by which epithelial cells transition into mesenchymal cells.
While the epithelial cells often stay bound together in layers, mesenchymal cells have gained motility and even stem-like properties\cite{nieto2016emt}.
Recent work has revealed the intriguing possibility that there are intermediate cell states along the EMT spectrum with each state having its own unique properties\cite{hong2015ovol2}.
Such intermediate states have the potential of being connected to tumor initiation, progression, and metastasis\cite{nieto2016emt}.
While there might be many other characteristics distinctive of mesenchymal cells that need further exploration, two qualities of mesenchymal cells are particularly relevant in the context of cancer and the immune system:
1) mesenchymal cells are less susceptible to immune clearance\cite{terry2017new} and
%
2) mesenchymal cells proliferate at a slower rate. \tcr{Ref}
 %A Adam said he had a reference in mind for this.

\tcr{Haven't addressed any of Qing's comments below this remark.}
The research on mesenchymal cells' immune evasiveness is quite conclusive with the research even going so far as establishing functional causes for this phenomenon.
As a cell is being targeted by cytotoxic immune cells for clearance, a physical connection between the two cells must be created in order to precipitate the lysing event.
This immunological synapse, however, is dependent upon the surface proteins of the target cell, and in the case of mesenchymal cells there is preliminary evidence that these surface markers are down-regulated in such a way that the immune synapse is more difficult to form\cite{terry2017new}.
There are also other structural changes that take place in mesenchymal cells that could contribute to this refractory nature towards immune clearance.
Regardless, the evidence currently demonstrates a ``slipperiness'' of mesenchymal cells in regards to immune lysing that in our model we will refer to as mesenchymal immune evasion, or MIE.
